---
title: "p8105_HW3_sp3804"
author: "Seonyoung Park (sp3804)"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(p8105.datasets)
library(tidyverse)
library(janitor)
```

```{R}

data("instacart")
glimpse(instacart)

````

Problem 0
This “problem” focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files.

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.
Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There ar ealso item varialbes -- name, aisle, department, and some numeric codes. 

How many aisles, and which ar emost items from?
```{r}
instacart %>%
  count(aisle)%>%
  arrange(desc(n))

```

Let's makd a plot

```{r}
instacart %>%
  count(aisle)%>%
  filter(n>10000)%>%
  ggplot(aes(x=aisle, y=n))+
  geom_point() # looks unclear 

## rotate the text in x-axis, reoder aisle in ascending order. 

instacart %>%
  count(aisle)%>%
  filter(n>10000)%>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x=aisle, y=n))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust =0.5, hjust=1))

```
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name)%>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank <4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()

```
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )

```

# Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
- pivot (day- numeric format? character? it affects to the third part in the third sub question)
```{r}



```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
--> group_by, summarise 
--> potential issue: day of week column (default; alphabetical order), we have to change the order

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
- scatterplot (x-axis minute time y; activities) + line 
- color ; day of week
- conclusion; weekday vs. weekend etc


#Problem 3
```{r}
library(p8105.datasets)
data("ny_noaa")

```

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
- data manipulation
- pplot - facet

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
- filtering
- show distribution
- plot