---
title: "p8105_HW3_sp3804"
author: "Seonyoung Park (sp3804)"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(p8105.datasets)
library(tidyverse)
library(janitor)
```

```{R}

data("instacart")
glimpse(instacart)

````

Problem 0
This “problem” focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files.

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.
Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There ar ealso item varialbes -- name, aisle, department, and some numeric codes. 

How many aisles, and which ar emost items from?
```{r}
instacart %>%
  count(aisle)%>%
  arrange(desc(n))

```

Let's makd a plot

```{r}
instacart %>%
  count(aisle)%>%
  filter(n>10000)%>%
  ggplot(aes(x=aisle, y=n))+
  geom_point() # looks unclear 

## rotate the text in x-axis, reoder aisle in ascending order. 

instacart %>%
  count(aisle)%>%
  filter(n>10000)%>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x=aisle, y=n))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust =0.5, hjust=1))

```
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name)%>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank <4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()

```
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )

```


